---
title: 'Titanic Dataset Inferences'
author: "Karthick Raja, Rahul Sathyajit"
output: html_document
---

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

#### Extracting the data and Function for the Hypothesis testing

get the data from here:
https://drive.google.com/file/d/1O3-thqx_6cyWQ6z7VxIsPwxfDfWTtzXZ/view?usp=sharing

``` {r}

#reading the data into a data frame
data1<-read.csv("C:\\Users\\Lenovo\\Desktop\\stats-assignment 2\\titanic.csv") #change the path here
#Test to be used
nptest<-function(x)
{
  Chisq<-chisq.test(x)
  Fis<-fisher.test(x)
  df<-data.frame(Method=c(Chisq$method, Fis$method),
                 P.Value=c(Chisq$p.value, Fis$p.value))
  df
}

sigtest<-function(x,y)
{
  kst<-ks.test(x,y)
  wil<-wilcox.test(x,y)
  df<-data.frame(Method=c(kst$method, wil$method),
                 P.Value=c(kst$p.value, wil$p.value))
  df
}

normtest<-function(x)
{
  library(nortest)
  s<-shapiro.test(x)
  ad<-ad.test(x)
  cvm<-cvm.test(x)
  ll<-lillie.test(x)
  sf<-sf.test(x)
  df<-data.frame(Method=c(s$method, ad$method, cvm$method, ll$method, sf$method),
                 P.Value=c(s$p.value, ad$p.value, cvm$p.value, ll$p.value, sf$p.value))
  df
}
``` 


####***1.Is there a significant difference in Age distribution between those who survived and those who did not?***

####Data Analysis

``` {r}

#having a table for age and survival
asu <- data.frame(data1$Age,data1$Survived)
#Dividing Survived and Not Survived
asu0<-subset(asu,asu$data1.Survived==0)
asu1<-subset(asu,asu$data1.Survived==1)
```

####Plots and Graphs

######Boxplot

``` {r}
boxplot(asu0$data1.Age,asu1$data1.Age,main='Not - Survived                       Survived')
```

######Histogram
``` {r}
hist(asu0$data1.Age,main = 'histogram of Age - Not Survived',xlab = 'Age')

```

``` {r}
hist(asu1$data1.Age,main = 'histogram of Age - Survived',xlab = 'Age')

```

######QQNorm - Plot
``` {r}
qqnorm(asu0$data1.Age,main = "QQ Normal plot for Age - Not Survived")

```
``` {r}
qqnorm(asu1$data1.Age,main = "QQ Normal plot for Age - Survived")
```

######Conducting Normality test for both the samples

``` {r}
normtest(asu0$data1.Age)
```

```{R}
normtest(asu1$data1.Age)
```

The P-Value of the Normality tests are less than 0.05 , strong evidence against null hypothesis ( The data follows normal distribution)

Both the data for survived and not survived people did not follow normal distribution

####Hypothesis Testing

***Null Hypothesis***: There is a no difference in Age distribution between those who survived and those who did not.

***Alternate Hypothesis***: There is a difference in Age Distribution between those who survived and those who did not

```{R}
#Test
wilcox.test(asu0$data1.Age,asu1$data1.Age)
```

>Results

The P-Value of the Wilcoxon rank sum test is 0.192. The is no strong evidence against Null hypothesis.

There is a **No significant difference ** between the age of the people who have survived and who have not. 

***
------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

####***2.Answer the same as above after controlling for Gender.***

####Data Analysis

```{R}
#Controlling for gender in the age table
gsu <- data.frame(data1$Gender,data1$Age,data1$Survived)
#Dividing Survived and Not Survived
gsu0<-subset(gsu,gsu$data1.Survived==0)
gsu1<-subset(gsu,gsu$data1.Survived==1)
```

####Graphs and Plots

######Barplot

```{R}
barplot(table(gsu0$data1.Gender),main='Barplot for Not Survived people',ylab='Number of people')
```


```{R}
barplot(table(gsu1$data1.Gender),main = 'Barplot for Survived People',ylab = 'Number of People')
```

```{R}

#Dividing the table for gender
msu0<-subset(gsu0,gsu0$data1.Gender=='male')
fsu0<-subset(gsu0,gsu0$data1.Gender=='female')
msu1<-subset(gsu1,gsu1$data1.Gender=='male')
fsu1<-subset(gsu1,gsu1$data1.Gender=='female')
```

######Box Plot
```{R}
boxplot(msu0$data1.Age,msu1$data1.Age,fsu0$data1.Age,fsu1$data1.Age,main='Male_not_survived     Male_Survived     Female_Not_Survived     Female_Survived',cex.main=0.9)
```

```{R out.extra='style="float:left"'}

hist(msu0$data1.Age,main = 'histogram of Age - Not Survived - Male',xlab = 'Age')

```
```{R}

hist(msu1$data1.Age,main = 'histogram of Age -  Survived - Male',xlab = 'Age')
```

```{R}
hist(fsu0$data1.Age,main = 'histogram of Age - Not Survived - Female',xlab = 'Age')
```

```{R}
hist(fsu1$data1.Age,main = 'histogram of Age - Survived - Female',xlab = 'Age')
```

######QQNorm Plot
```{R}
qqnorm(msu0$data1.Age,main = "QQ Normal plot for Age - Not Survived - Male")

```

```{R}
qqnorm(msu1$data1.Age,main = "QQ Normal plot for Age -  Survived - Male")
```

```{R}
qqnorm(fsu0$data1.Age,main = "QQ Normal plot for Age - Not Survived - Female")
```

```{R}
qqnorm(fsu1$data1.Age,main = "QQ Normal plot for Age -  Survived - Female")
```
Normality test for both the samples

```{R}
normtest(msu0$data1.Age)
```

```{R}
normtest(msu1$data1.Age)
```

```{R}
normtest(fsu0$data1.Age)
```

```{R}
normtest(fsu1$data1.Age)
```

######Normality Test Results

Female Passengers who did not survive have a normally distributed data. 

All the other categories like Male survived, Male not survived, Female survived data are not normal.

####Hypothesis Testing - For Male

***NUll Hypothesis: *** There is no signigicant difference in Age distribution between those who survived and those who did not for male.

***Alternative Hypothesis: ***  There is signigicant difference in Age distribution between those who survived and those who did not for male.

```{R}
#Tests
sigtest(msu0$data1.Age,msu1$data1.Age)
```

>Results

The P-Value of both the two-sample KS test and Wilcoxon rank sum test were found to be less than 0.05.

There is a strong evidance against the Null Hypothesis.

There is signigicant difference in Age distribution between those who survived and those who did not for male.

####Hypothesis Testing - For Female

***NUll Hypothesis: *** There is no signigicant difference in Age distribution between those who survived and those who did not for Female

***Alternative Hypothesis: ***  There is signigicant difference in Age distribution between those who survived and those who did not for Female

```{R}
#Tests
sigtest(fsu0$data1.Age,fsu1$data1.Age)
```

>Results

The P-Value of both the two-sample KS test and Wilcoxon rank sum test were found to be less than 0.05.

There is a strong evidance against the Null Hypothesis.

There is signigicant difference in Age distribution between those who survived and those who did not for Female.

***
--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

####***3. Remark on how Age affected the Survival Probability of a passenger on board the Titanic, based on consolidations of your findings in 1 and 2 above***
   
####Data Analysis

```{R}
#copy to staging table
asuc1<-asu1
asuc0<-asu0
asuc0<-cut(asuc0$data1.Age,c(0,18,30,55,150))
asuc1<-cut(asuc1$data1.Age,c(0,18,30,55,150))

x<-table(asuc1)
y<-table(asuc0)

x<-data.frame(t(t(table(asuc1))))
y<-data.frame(t(t(table(asuc0))))
x<-x[-2]
y<-y[-2]
#combining tables
AgeTable<-cbind(x,y$Freq)

#renaming columns
names(AgeTable)[names(AgeTable) == "asuc1"] <- "def"
names(AgeTable)[names(AgeTable) == "Freq"] <- "Survived"
names(AgeTable)[names(AgeTable) == "y$Freq"] <- "Un-Survived"
rownames(AgeTable)<-c("Young","Young Adult","Adult","Old")
AgeTable
```

####Plots and Graphs

######Barplots

```{R}
#BarPlot

temp<-AgeTable$Survived
names(temp)<-c("young","young adult","Adult","old")
barplot(temp,main = 'Survived')

```
```{r}
boxplot(temp, main='Boxplot for Survived')
```

```{r}
temp<-AgeTable$`Un-Survived`
names(temp)<-c("young","young adult","Adult","old")
barplot(temp,main='Un-Survived')
```
```{r}
boxplot(temp, main='Boxplot for Not-Survived')
```

####Hypothesis Testing

***Null Hypothesis:*** The Survival Probablity is not affected by the Age of the passanger
The survival probability is homogenous across all the Age buckets

***Alternate Hypothesis:*** The Survival Probability is affected by the Age of the Passanger
The Survival Probability is Not Homogenous across all the Age buckets

```{r}
AgeTable<-AgeTable[,-1]
chisq.test(AgeTable)
```

>Results

The p value of the chi-quare test of Homogeneity is 2.2e-16, which is less than 0.05.

There is a strong evidence against the null hypothesis

The Survival Probability is affected by the Age Distribution.

***
-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

####***4. Is there a significant difference in Survival Probability between the two genders?***

####Data Analysis

#####Gender based Survival Table

```{r}
#copy to staging table
gen<-data.frame(data1$Gender,data1$Survived)
gen0<-subset(gen,gen$data1.Survived==0)
gen1<-subset(gen,gen$data1.Survived==1)

gen0<-data.frame(t(t(table(gen0$data1.Gender))))

gen1<-data.frame(t(t(table(gen1$data1.Gender))))

gen0<-gen0[-2]
gen0<-gen0[-1]
gen1<-gen1[-2]
gen1<-gen1[-1]
#Combining Table
GenderTable<-cbind(gen0,gen1$Freq)
rownames(GenderTable)<-c("Female","Male")
colnames(GenderTable)<-c("Survived","Not-Survived")

GenderTable
```

```{r}
temp<-GenderTable$Survived
names(temp)<-c("Female","Male")
barplot(temp,main = "Survived")
```

```{r}
temp<-GenderTable$`Not-Survived`
names(temp)<-c("Female","Male")
barplot(temp,main = "Not-Survived")
```

####Hypothesis Testing

***Null Hypothesis:*** There is no difference in the survival probability between the two genders

***Alternative Hypothesis:*** There is a difference in the survival probability between the two genders

```{r}
nptest(GenderTable)
```
>Results

Both the Chi-Sq test and Fisher test are giving P-Value less than 0.05.

There is a strong evidence against Null Hypothesis

The Survival Rate is dependent on the Gender based on chi square test for independence and Fisher's Exact 

***
--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

####***5. Is there a significant difference in Survival Probability between the three passenger classes?***
  
####Data Analysis

#####Passanger Class Table
```{r}

temp<-subset(data1,data1$Survived==0)
temp<-subset(temp,temp$PClass=='1st')
class1n<-data.frame(temp$PClass,temp$Survived)

temp<-subset(data1,data1$Survived==1)
temp<-subset(temp,temp$PClass=='1st')
class1s<-data.frame(temp$PClass,temp$Survived)

temp<-subset(data1,data1$Survived==0)
temp<-subset(temp,temp$PClass=='2nd')
class2n<-data.frame(temp$PClass,temp$Survived)

temp<-subset(data1,data1$Survived==1)
temp<-subset(temp,temp$PClass=='2nd')
class2s<-data.frame(temp$PClass,temp$Survived)

temp<-subset(data1,data1$Survived==0)
temp<-subset(temp,temp$PClass=='3rd')
class3n<-data.frame(temp$PClass,temp$Survived)

temp<-subset(data1,data1$Survived==1)
temp<-subset(temp,temp$PClass=='3rd')
class3s<-data.frame(temp$PClass,temp$Survived)

class1s<-table(class1s)
class1n<-table(class1n)
class2s<-table(class2s)
class2n<-table(class2n)
class3n<-table(class3n)
class3s<-table(class3s)

ClassTable<-matrix(c(class1s[-2:-3],class1n[-2:-3],class2s[2],class2n[2],class3s[3],class3n[3]),ncol=2,byrow=TRUE)
rownames(ClassTable)<-c("Class-1","Class-2","Class-3")
colnames(ClassTable)<-c("Survived","Un-Survived")
ClassTable
```
####Hypothesis Testing

***Null Hypothesis:*** The Survival probability is same across all the passanger class

***Alternate Hypothesis:*** The Survival probability is not same across all the passanger class


```{r}
nptest(ClassTable)
```

>Results
The P-Value given by both Chi-Sq test and Fisher's Exact test are less than 0.05. 

There is a strong evidence against Null hypothesis.

The survival probability is not distributed evenly across the P-Class.

***
--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

####***6. Is there a significant difference in Survival Probability between the two genders even after taking the effect of Passenger Class into Account?***

####Data Analysis
   
```{r}
#The Gender Table Generator for Different classes
gtrans<-function(x)
{
gen<-data.frame(q6$Gender,q6$Survived)
gen0<-subset(gen,gen$q6.Survived==0)
gen1<-subset(gen,gen$q6.Survived==1)

gen0<-data.frame(t(t(table(gen0$q6.Gender))))

gen1<-data.frame(t(t(table(gen1$q6.Gender))))

gen0<-gen0[-2]
gen0<-gen0[-1]
gen1<-gen1[-2]
gen1<-gen1[-1]
#Combining Table
GenderTable<-cbind(gen0,gen1$Freq)
rownames(GenderTable)<-c("Female","Male")
colnames(GenderTable)<-c("Survived","Not-Survived")

return(GenderTable)
}
```

####Table for Class 1
```{r}
q6<-subset(data1,data1$PClass=='1st')
temp<-gtrans(q6)
temp
```
####Hypothesis Testing for Class-1 Passengers

***Null Hypothesis:*** There is no difference between the survival proportion between the male and female passanger of class 1

***Alternate Hypothesis:*** There is a significant difference between the survival proportion between the male and female passanger of class 1

```{r}
nptest(temp)
```
>Results

The P-Value of both the Chi-Sq test and Fisher Exact Test is less than 0.05

There is a strong evidence against Null Hypothesis

Theere is a significant difference between the Survival Probablity of Male and Female of the class-1 Passangers. 

####Table for Class 2
```{r}
q6<-subset(data1,data1$PClass=='2nd')
temp<-gtrans(q6)
temp
```
####Hypothesis Testing for Class-2 Passengers

***Null Hypothesis:*** There is no difference between the survival proportion between the male and female passanger of class 2

***Alternate Hypothesis:*** There is a significant difference between the survival proportion between the male and female passanger of class 2

```{r}
nptest(temp)
```
>Results

The P-Value of both the Chi-Sq test and Fisher Exact Test is less than 0.05

There is a strong evidence against Null Hypothesis

Theere is a significant difference between the Survival Probablity of Male and Female of the class-2 Passangers.

####Table for Class-3

```{r}
q6<-subset(data1,data1$PClass=='3rd')
temp<-gtrans(q6)
temp
```
####Hypothesis Testing for Class-3 Passengers

***Null Hypothesis:*** There is no difference between the survival proportion between the male and female passanger of class 3

***Alternate Hypothesis:*** There is a significant difference between the survival proportion between the male and female passanger of class 3

```{r}
chisq.test(temp)
```

>Results

The P-Value of both the Chi-Sq test and Fisher Exact Test is less than 0.05

There is a strong evidence against Null Hypothesis

There is a significant difference between the Survival Probablity of Male and Female of the class-3 Passangers.
