---
title: "SVM and KNN"
author: "Karthick"
date: "31 March 2018"
output: html_document
---

####7.1 One Hot Encoding

The reason for running the SVM and KNN together is that both the algorithms can't tolerate the presence of a categorical variables in the predictors. It is necessary for us to convert ( one hot encode ) the categorical variables into numerical variables before sending them to train.


***Encoding Process***

1. Label the numerical data with different numbers

2. create a separate column for each variable containing either 0 or 1, indicating the presence or absence of the variable

3. columns are created such the variable having 3 levels are represented by 2 columns with the following combinations 1: (0,0) level 1 (1,0) level 2 (0,1) level 2

```{r }
# Importing the dataset

# Importing the dataset
dfFinal <- read.csv("dfFinal.csv")
dfFinal$totDep <- as.factor(dfFinal$totDep)
dfFinal$PHON <- as.factor(dfFinal$PHON)
dfFinal$bad <- as.factor(dfFinal$bad)

dataset = dfFinal


# Encoding categorical data
for(unique_value in unique(dataset$totDep)){
  
  
  dataset[paste("totDep", unique_value, sep = ".")] <- ifelse(dataset$totDep == unique_value, 1, 0)
}
dataset$totDep <- NULL
dataset$totDep.5 <-NULL

dataset$RES = factor(dataset$RES,
                     levels = c('F', 'N', 'O','P','U'),
                     labels = c(1, 2, 3,4,5))

for(unique_value in unique(dataset$RES)){
  
  
  dataset[paste("RES", unique_value, sep = ".")] <- ifelse(dataset$RES == unique_value, 1, 0)
}
dataset$RES <- NULL
dataset$RES.5 <-NULL


dataset$AES = factor(dataset$AES,
                     levels = c('B', 'E', 'M', 'N', 'P', 'R', 'T', 'U' ,'V', 'W' ,'Z'),
                     labels = c(1,2,3,4,5,6,7,8,9,10,11))

for(unique_value in unique(dataset$AES)){
  
  
  dataset[paste("AES", unique_value, sep = ".")] <- ifelse(dataset$AES == unique_value, 1, 0)
}
dataset$AES <- NULL
dataset$AES.11 <-NULL

dataset$PHON = factor(dataset$PHON,
                      levels = c('0','1'),
                      labels = c(1,2))

for(unique_value in unique(dataset$PHON)){
  
  
  dataset[paste("PHON", unique_value, sep = ".")] <- ifelse(dataset$PHON == unique_value, 1, 0)
}
dataset$PHON <- NULL
dataset$PHON.2 <-NULL



```


####7.2 SVM model

```{r }

# Splitting the dataset into the Training set and Test set
# install.packages('caTools')
library(caTools)

split = sample.split(dataset$bad, SplitRatio = 0.75)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)

# Fitting Kernel SVM to the Training set
# install.packages('e1071')
library(e1071)
classifierSVM = svm(formula = bad ~.,
                 data = training_set,
                 cost = 10,
                 type = 'C-classification',
                 kernel = 'sigmoid')

# Predicting the Test set results
y_pred_svm = predict(classifierSVM, newdata = test_set[-4])

# Making the Confusion Matrix
cm = table(test_set[, 4], y_pred_svm)
cm


```
The model is trained with the sigmoid function., a special case of logistic function.

```{r }
accuracy = (cm[1,1] + cm[2,2]) / (cm[1,1] + cm[2,2] + cm[1,2] + cm[2,1])
sensitivity=(cm[2,2]/(cm[2,2]+cm[2,1]))
specificity=(cm[1,1]/(cm[1,1]+cm[1,2]))

print("the accuracy of the model is ")
print(accuracy)

print("the sensitivity of the model is ")
print(sensitivity)

print("the specificity of the model is ")
print(specificity)

```

####7.3 KNN
